---
title: "Modeling and Forecasting Daily Electricity Load"
subtitle: "IDI II"
author: "Pablo Benavides-Herrera"
date: "2019-11-15"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
    highlight: tango
---

We use the data provided by CENACE regarding the daily electricity load for the western regional control management (GCROC,for its initials in Spanish).

# Initial setup

Here we load the required packages, define functions and load and wrangle the data.

## Packages

```{r pkgs, message=FALSE}
library(tidyverse) # call to the tidyverse packages
library(lubridate) # tidy date manipulation
# library(pracma) # 
# library(xtable)
library(plotly) # interactive plots
library(patchwork) # plots arrangement
library(tsibble) # tidy time series tibbles
library(fable) # time series tidy modeling and forecasting
library(feasts)
# library(cowplot)
# library(withr) # change global options, locally
# library(cluster) # for cluster analysis
# library(factoextra) # cluster visualization
```

## Functions
Defining a MAPE function to be used.

```{r fun - mape}
mape <- function(y=consumo,y_hat){
  mape = abs((y-y_hat)/y)*100
  return(mape)
}
```

A function to filter data on ggplot layers.

```{r fun - pick}
pick <- function(condition){
  function(d) d %>% filter(condition)
}
```


Descriptive statistics function.
```{r fun stats}
d_stats <- function(x){
  list(media = mean(x,na.rm = T),
       `desv. std` = sd(x,na.rm = T),
       min = min(x,na.rm = T),
       mediana = median(x,na.rm = T),
       max = max(x,na.rm = T),
       n = length(x),
       NAs = sum(is.na(x)))
}
```


## The data

The main data to be used is the daily load. We also use holidays data as to get insights on the variation.

```{r data - load, message=FALSE}
datos <- read_csv("gcroc.csv")
festivos <- read_csv("festivos_utf8.csv")
datos <- datos %>% 
  left_join(festivos, by = c("fecha" = "fecha"))

datos <- datos %>% 
  mutate(fecha = dmy(fecha),
         dia_semana = factor(dia_semana,
                             levels = c(1:7)),
         Day_type = case_when(
           festivo %in% festivos$festivo ~ "Holiday",
           wday(fecha,week_start = 1) == 1 ~ "Monday",
           wday(fecha, week_start = 1) %in% 1:5 ~ "Weekday",
           TRUE ~ "Weekend"
         ) %>% factor(levels = c("Monday","Weekday",
                                 "Weekend","Holiday"))
         ) %>% 
  as_tsibble(index = fecha) %>% 
  rename(`temp min` = temp_min, 
         `temp max` = temp_max,
         `temp prom` = temp_prom)
datos
```



```{r plot, fig.width=12, fig.height=10}
g1 <- ggplot(datos,aes(x = fecha, y = MWh))
# Time plot
p1 <- g1 + geom_line() + 
  ggtitle("Historic Load") 
# Day type boxplot
p2 <- g1 + geom_boxplot(aes(x = Day_type, color = Day_type)) + 
  theme(legend.position = "none",
        axis.text.x =  element_text(angle = 45,
                                      hjust = 1)) + 
  ggtitle("Load Across Day Types")
# Month boxplot
p3 <- g1 + geom_boxplot(aes(x = month(fecha, label = T, abbr = T), 
  color = month(fecha, label = T, abbr = T)
  )
  ) + 
  theme(legend.position = "none",
        axis.text.x =  element_text(angle = 45,
                                      hjust = 1)) + 
  ggtitle("Load Across Months")
# plot group 1
pg1 <- p1 / (p2 | p3)

pg1 + plot_annotation(
  title = " Daily Electricity Load in the GCROC"
)
# TS Decomposition
datos %>% 
  model(STL(MWh)) %>% 
  components() %>% 
  autoplot()
```

```{r}
# descr_stats <- datos %>% 
#   as_tibble() %>% 
#   select_if(is_double) %>% 
#   sapply(d_stats)
# 
# descr_stats

datos  %>%  
  features_if(is_double,features =
                         list(media = mean,
       `desv. std` = sd,
       min = min,
       mediana = median,
       max = max,
       n = length), na.rm = T
       ) %>% 
  pivot_longer(everything()) %>% 
  separate(name, into = c("var","stat"), sep = "_") %>% 
  pivot_wider(names_from = "var",values_from = "value")
```

```{r}
datos %>% 
  features(log(MWh), unitroot_nsdiffs)

datos %>% 
  features(difference(log(MWh),7), unitroot_ndiffs)

datos %>% 
  gg_tsdisplay(MWh %>% log() %>%  difference(7) %>% difference(), plot_type = "partial", lag_max = 49)
```







