---
title: "Modeling and Forecasting Daily Electricity Load"
subtitle: "IDI II"
author: "Pablo Benavides-Herrera"
date: "2019-11-15"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
    theme: cerulean
    highlight: tango
---

We use the data provided by CENACE regarding the daily electricity load for the western regional control management (GCROC,for its initials in Spanish).

# Initial setup

Here we load the required packages, define functions and load and wrangle the data.

## Packages

```{r pkgs, message=FALSE}
library(tidyverse) # call to the tidyverse packages
library(lubridate) # tidy date manipulation
# library(pracma) # 
library(xtable)
library(plotly) # interactive plots
library(patchwork) # plots arrangement
library(tsibble) # tidy time series tibbles
library(fable) # time series tidy modeling and forecasting
library(feasts)
library(fable.prophet)
# library(cowplot)
# library(withr) # change global options, locally
# library(cluster) # for cluster analysis
# library(factoextra) # cluster visualization
```

## Functions
Defining a MAPE function to be used.

```{r fun - mape}
mape <- function(y=consumo,y_hat){
  mape = abs((y-y_hat)/y)*100
  return(mape)
}
```

A function to filter data on ggplot layers.

```{r fun - pick}
pick <- function(condition){
  function(d) d %>% filter(condition)
}
```


Descriptive statistics function.
```{r fun stats}
# d_stats <- function(x){
#   list(media = mean(x,na.rm = T),
#        `desv. std` = sd(x,na.rm = T),
#        min = min(x,na.rm = T),
#        mediana = median(x,na.rm = T),
#        max = max(x,na.rm = T),
#        n = length(x),
#        NAs = sum(is.na(x)))
# }
```


## The data

The main data to be used is the daily load. We also use holidays data as to get insights on the variation.

```{r data - load, message=FALSE}
datos <- read_csv("gcroc.csv")
festivos <- read_csv("festivos_utf8.csv")
datos <- datos %>% 
  left_join(festivos, by = c("fecha" = "fecha"))

datos <- datos %>% 
  mutate(fecha = dmy(fecha),
         dia_semana = factor(dia_semana,
                             levels = c(1:7)),
         Day_type = case_when(
           festivo %in% festivos$festivo ~ "Holiday",
           wday(fecha,week_start = 1) == 1 ~ "Monday",
           wday(fecha, week_start = 1) %in% 1:5 ~ "Weekday",
           TRUE ~ "Weekend"
         ) %>% factor(levels = c("Monday","Weekday",
                                 "Weekend","Holiday")),
         Month = month(fecha, label = TRUE) %>% factor(ordered = F)
         ) %>% 
  as_tsibble(index = fecha) %>% 
  rename(`temp min` = temp_min, 
         `temp max` = temp_max,
         `temp prom` = temp_prom)
datos
```

```{r, include=FALSE}
# To print in LaTeX form
datos %>% 
  as_tibble() %>%  
  mutate(fecha = as.character(fecha)) %>%  slice_head(n = 10) %>% 
  xtable() %>% print()
```




# Time Series Analysis

## Descriptive Statistics

```{r plot, fig.width=12, fig.height=10}
g1 <- ggplot(datos,aes(x = fecha, y = MWh))
# Time plot
p1 <- g1 + geom_line() + 
  ggtitle("Historic Load") 
# Day type boxplot
p2 <- g1 + geom_boxplot(aes(x = Day_type, color = Day_type)) + 
  theme(legend.position = "none",
        axis.text.x =  element_text(angle = 45,
                                      hjust = 1)) + 
  ggtitle("Load Across Day Types")
# Month boxplot
p3 <- g1 + geom_boxplot(aes(x = Month, 
  color = Month)
  ) + 
  theme(legend.position = "none",
        axis.text.x =  element_text(angle = 45,
                                      hjust = 1)) + 
  ggtitle("Load Across Months")
# plot group 1
pg1 <- p1 / (p2 | p3)

pg1 + plot_annotation(
  title = " Daily Electricity Load in the GCROC"
)
```

We see differences across day types and across months, suggesting year seasonality.

```{r stats}
datos  %>%  
  features_if(is_double,features =
                         list(media = mean,
       `desv. std` = sd,
       min = min,
       mediana = median,
       max = max,
       n = length), na.rm = T
       ) %>% 
  pivot_longer(everything()) %>% 
  separate(name, into = c("var","stat"), sep = "_") %>% 
  pivot_wider(names_from = "var",values_from = "value")
```

We check for the relationship between electricity load and temperature.

```{r MWh v. temp, fig.width=12, fig.height=10}
p4 <- datos %>% 
  ggplot(aes(x = `temp max`, y = MWh)) + 
  geom_point() + ggtitle("Maximum Temperature") + 
  xlab("(C°)") + geom_smooth(method = "lm", formula = y ~ x + I(x^2))# +
  # facet_wrap(~ factor(year(fecha)))

p5 <- datos %>% 
  ggplot(aes(x = `temp min`, y = MWh)) + 
  geom_point() + ggtitle("Minimum temperature") + 
  xlab("(C°)") + theme(legend.position = "none") + geom_smooth(method = "lm", formula = y ~ x + I(x^2))# +
  # facet_wrap(~ factor(year(fecha)))

pg2 <- p4 / p5

pg2 + plot_annotation(title = "Electricity Load as a Function of Temperature")

p4
```


## Decomposition Analysis

```{r TS Decomposition}
datos %>% 
  model(STL(log(MWh))) %>% 
  components() %>% 
  autoplot()
# Components
datos %>% 
  model(STL(log(MWh))) %>% 
  components()
# STL features
datos %>% 
  features(MWh, feat_stl)
```




## Fourier Analysis

```{r manual fourier}
datos_components_log <- datos %>% 
  model(STL(log(MWh))) %>% 
  components() %>% 
  mutate(detrended = `log(MWh)` - trend)

autoplot(datos_components_log, detrended)

datos_acf <- datos_components_log %>% 
  ACF(detrended, lag_max = 1850) %>% 
  select(acf) 

datos_acf <- datos_acf %>% 
  mutate(fourier = abs(fft(datos_acf$acf)^2)
         )
autoplot(datos_acf,fourier)

periodos <- datos_acf %>% 
  top_n(fourier,n = 10)

coef_fourier_sin <- t(sin(2 * pi / t(t(periodos$fourier)) %*% t(seq_along(datos$fecha))))
coef_fourier_cos <- t(cos(2 * pi / t(t(periodos$fourier)) %*% t(seq_along(datos$fecha))))
a <- tibble(coef_fourier_sin,coef_fourier_cos)
```

# TS Models

## Fitting the models

```{r model}
h <- 365
train <- datos %>% 
  slice(1 : (n() - h))
fit <- train %>% 
  model(
    # `LM + dummies` = TSLM(log(MWh) ~ Day_type + Month),
    # `LM + dummies + season` = TSLM(log(MWh) ~ Day_type + Month + season()),
    # `LM + dummies + max temp` = TSLM(log(MWh) ~ Day_type + Month + `temp max` + I(`temp max`^2)),
    # `LM + dummies + season + max temp` = TSLM(log(MWh) ~ Day_type + Month + `temp max` + I(`temp max`^2) + season()),
    # `LM + ARIMA errors` = ARIMA(log(MWh) ~ Day_type + Month + `temp max` + I(`temp max`^2) + pdq(3,1,2) + PDQ(0,0,0)),
    # `LM + season + ARIMA errors` = ARIMA(log(MWh) ~ Day_type + Month + `temp max` + I(`temp max`^2) + season() + PDQ(0,0,0)),
    `Harmonic reg + dummies` = ARIMA(log(MWh) ~ Day_type +
                      fourier(period = "1 year", K = 20) + 
                      fourier(period = "1 week", K = 2) + PDQ(0,0,0)) 
    )
report(fit)
# report(fit %>% select(7))
# tidy(fit)
```

```{r}
train %>% 
  features(log(MWh),unitroot_nsdiffs)
train %>% 
  features(log(MWh) %>% difference(7),unitroot_ndiffs)
```
```{r}
h <- 365
train <- datos %>% 
  slice(1 : (n() - h))
```

```{r}
lambda <- train %>% 
  features(MWh, features = guerrero) %>% 
  pull(lambda_guerrero)
lambda
```


Implementing different models:

```{r}
fit2 <- train %>% 
  model(
    ETS = ETS(box_cox(MWh, lambda)),
    ARIMA = ARIMA(box_cox(MWh, lambda)),
    Harmonic = ARIMA(box_cox(MWh, lambda) ~ fourier(period = "1 year", K = 20) + 
                       fourier(period = "1 week", K = 2) + 
                       PDQ(0,0,0)),
    # `LM + ARIMA errors` = ARIMA(box_cox(MWh, lambda) ~ Day_type + Month + `temp max` + I(`temp max`^2) + pdq(3,1,2) + PDQ(0,0,0)),
    # # `LM + season + ARIMA errors` = ARIMA(log(MWh) ~ Day_type + Month + `temp max` + I(`temp max`^2) + season() + PDQ(0,0,0)),
    # `Harmonic reg + dummies` = ARIMA(box_cox(MWh, lambda) ~ Day_type +
    #                   fourier(period = "1 year", K = 20) + 
    #                   fourier(period = "1 week", K = 2) + PDQ(0,0,0)),
    Prophet = prophet(box_cox(MWh))
    )
report(fit2)
```



## Residuals Diagnostics

```{r resids}
fit %>% 
  # select(6) %>% 
  gg_tsresiduals()
```

## Backtesting

```{r backtesting, fig.width=12, fig.height=10}
test <- datos %>% 
  slice((n()-h+1):n())
  # new_data(train, 365) %>%
  # mutate(Day_type = slice(datos,(n()-364):n()) %>% as_tibble() %>%  select(Day_type) %>% as_tibble() %>% c(),
  #        Month = slice(datos,(n()-364):n()) %>% as_tibble() %>% select(Month) %>% c(),
  #        `temp max` = slice(datos,(n()-364):n()) %>% as_tibble() %>% select(`temp max`) %>% c()
  #        )
fcst <- forecast(fit, test)
p6 <- fcst %>% 
  autoplot(datos, level = NULL) + ylab("Electricity demand (MWh)") +
  facet_wrap(~.model) + 
  tidyquant::coord_x_date(xlim = c("2018-09-01","2019-09-22")
                          ) + theme(legend.position = "none") 
p7 <- fcst %>% 
  filter(str_detect(.model, "ARIMA")) %>% 
  autoplot(datos, level = NULL) + ylab("Electricity demand (MWh)") +
  facet_wrap(~.model) + 
  tidyquant::coord_x_date(xlim = c("2018-09-01","2019-09-22"), ylim = c(120000,220000)) + theme(legend.position = "none")
p6 / p7
```

The models' performance.

```{r accuracy}
accuracy(fcst, datos) %>% 
  arrange(MAPE)
```


```{r}
prueba <- train %>% slice(1:5)
prueba
prueba %>% new_data(5)

datos %>% 
  select(fecha) %>%
  slice(n()) - 365
```

